---
title: "Assignment 4"
output:
  html_document: default
  pdf_document: default
  word_document: default
date: "2023-10-23"
---
Install required packages
```{r}
library(tidyverse)  # data manipulation
# install.packages("factoextra") 
library(factoextra) 
library(ISLR)
library(httr)
library(flexclust)
library(caret)
set.seed(123)
```
Import data set
```{r}
library(readr)
Pharmaceuticals <- read_csv("Pharmaceuticals.csv")
View(Pharmaceuticals)
```
## a. Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on.
```{r}
# Select variables 1 to 9
Pharmaceuticals.data <- Pharmaceuticals[, -c(1,2,12,13,14)]

# Scale the data
Pharmaceuticals.scaled<-scale(Pharmaceuticals.data)
distance <-get_dist(Pharmaceuticals.scaled)
fviz_dist(distance)

# Determine k 
fviz_nbclust(Pharmaceuticals.scaled, kmeans, method = "wss") # using Elbow Method

fviz_nbclust(Pharmaceuticals.scaled, kmeans, method = "silhouette") # Using silhouette method
```
From analyzing the graphs we can see the best value of k is 5. Adding more or having less clusters than 5 will bring less improvement to cluster homogeneity.
```{r}
# Cluster analysis - we chose k-means since we know the number of clusters that are best for the analysis
# Numer of clusters formed = 5. We can find by using an elbow chart and the Silhouette Method
# By default we use Euclidean distance
k5 <- kmeans(Pharmaceuticals.scaled, centers = 5, nstart = 25)

# Visualize the output
k5$centers
k5$size
fviz_cluster(k5, data = Pharmaceuticals.scaled)
```
## b. Interpret the clusters with respect to the numerical variables used in forming the clusters.
```{r}
# Print the mean value of the variables by cluster
Pharmaceuticals.data %>%
  mutate(Cluster = k5$cluster) %>%
  group_by(Cluster) %>%
  summarise_all("mean")

# Print all data adding a cluster column. To visualize the cluster assigned to each company.
full.data <- cbind(Pharmaceuticals, cluster = k5$cluster)
tibble(full.data)
```
## c. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)
```{r}
# Breakdown clusters by median recommendation
Recomendation <- table(k5$cluster, Pharmaceuticals$Median_Recommendation) 
names(dimnames(Recomendation)) <- c("Cluster", "Recommendation")
Recomendation <- addmargins(Recomendation)
Recomendation

# Breakdown cluster by the location of the firm's headquarters
Location.firm <- table(k5$cluster, Pharmaceuticals$Location)
names(dimnames(Location.firm)) <- c("Cluster", "Location")
Location.firm <- addmargins(Location.firm)
Location.firm

# Breakdown clusters by the stock exchange on which the firm is listed
Stock.Exchange <- table(k5$cluster, Pharmaceuticals$Exchange)
names(dimnames(Stock.Exchange)) <- c("Cluster", "Stock Exchange")
Stock.Exchange <- addmargins(Stock.Exchange)
Stock.Exchange
```
```{r}
# Create a new data set to include the cluster column
Pharma.Cluster<-Pharmaceuticals
Pharma.Cluster$Cluster <- as.factor(k5$cluster)

# To create a mode table we define the mode function
mode_stat <- function(x) {
  tbl <- table(x)
  name <- names(tbl)[which.max(tbl)]
  if (is.null(name)) {
    return(NA)
  } else {
    return(name)
  }
}

pattern.table <- Pharma.Cluster[,c(12:15)]
pattern.table <- aggregate(pattern.table[-4], pattern.table[4], mode_stat)

print(pattern.table)
```
## d. Provide an appropriate name for each cluster using any or all of the variables in the dataset.

Cluster 1 - Moderate Risk High PE Ratio - Moderate asset turnover, low profit margin, recommendation to hold
Cluster 2 - High Company Growth High Risk - moderate net profit margin, high revenue growth, moderate recommendation to buy 
Cluster 3 - Moderate Risk High Profitability - high net profit margin, recommended to hold
Cluster 4 - High Risk Low Profitability - low ROA, high leverage, recommended to hold
Cluster 5 - Stable High Profit - high asset turnover, recommended to hold
